SU($n$) ferromagnets near the Heisenberg point

### 2020-06-17 collective_sun/sun_perturbation.tex
amended some discussions about the operator content of energy eigenstates

In the case of an SU($n$) spin system, we can similarly start with the
ground-state PS manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited
states by the action of multi-body operators, $\O\p{\m w,X}\ket{m}$.
The choice of tensors $\m w$ for generating states in the $r$-magnon
manifold $\M_r$ is the same as before, but we now have the additional
complication of having to choose $r$-spin operators, $X$.  As with the
choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose
tensor products of {\it diagonal} single-spin operators (that is,
diagonal in a fixed single-particle basis), but there will now be
$\sim\p{n-1}^r$ choices: one of $n-1$ diagonal operators (that are
linearly independent of the identity) for each of $r$ tensor factors.
We discuss the exact choice of operators in more detail in Appendix
\ref{sec:operator_content}.  To summarize, we define $n$ single-spin
operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$, for
which $Z_\nu=\pm\sigma_\z$ in the case of SU(2), and we choose a list
of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
$\sum_\mu z_\mu=r$ to define the $r$-spin operator
\begin{align}
  X_z \equiv \bigotimes_{\nu=1}^n Z_\nu^{\otimes z_\nu}
  = \underbrace{Z_1 \otimes \cdots \otimes Z_1}_{z_1~\t{times}}
  \otimes
  \underbrace{Z_2 \otimes \cdots \otimes Z_2}_{z_2~\t{times}}
  \otimes \cdots \otimes
  \underbrace{Z_n \otimes \cdots \otimes Z_n}_{z_n~\t{times}}.
\end{align}
Our starting basis for $\M_r$ then consists of states
$\ket{m,\m w,z}\equiv\O\p{\m w,X_z}\ket{m}$, where we enforce that
$z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number
of the single-spin state $\nu$), and that $z$ is zero in an entry
where $m$ is maximal: $z_\nu=0$ for
$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  This second restriction is
equivalent to removing an operator from $\set{Z_\nu:\nu\in\ZZ_n}$,
making the remaining set linearly independent of the identity
operator.  The basis $\set{\ket{m,\m w,z}}$ for $\M_r$ is generally
complete, but {\it not} orthonormal, so as a final step we construct
an orthonormal basis $\set{\ket{m,\m w,X}}$, where for operators $X$
we define $\ket{m,\m w,X}\equiv\O\p{\m w,X}\ket{m}$ (see Appendix
\ref{sec:operator_content}).

@@

The above restrictions on $z$ do not {\it guarantee} that
$\ket{m,\m w,z}\ne0$, or that the restricted basis
$\set{\ket{m,\m w,z}}$ for $\M_r$ will not be overcomplete.
Furthermore, this restricted basis is not orthogonal, which is to say
that generally $\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.  In
order to construct an orthonormal basis for $\M_r$, we therefore need
to build a matrix $B_m$ with matrix elements
\begin{align}
  B_m\p{\m v,y;\m w,z} \equiv \bk{m,\m v,y|m,\m w,z},
\end{align}
where $y,z$ satisfy the restrictions outlined above, and $\m v,\m w$
are chosen from the set of tensors that generate eigenstates of $H_0$.
Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form
$\sum_{\m w,z}c_{\m w z}\ket{m,\m w,z}$, whose squared norms are equal
to their corresponding eigenvalues.  Note that the matrix $B_m$ has a
block-diagonal structure, as it cannot (for example) couple states
$\ket{m,\m v,y},\ket{m,\m w,z}$ whose tensors $\m v,\m w$ generate
states with different interaction energies.  We therefore only need to
build and diagonalize this matrix independently within each block.

>>>>

In the case of an SU($n$) spin system, we can similarly start with the
ground-state PS manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited
states by the action of multi-body operators, $\O\p{\m w,X}\ket{m}$.
The choice of tensors $\m w$ for generating states in the $r$-magnon
manifold $\M_r$ is the same as before, but we now have the additional
complication of having to choose $r$-spin operators, $X$.  As with the
choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose
tensor products of {\it diagonal} single-spin operators (that is,
diagonal in a fixed single-particle basis), but there will now be
$\sim\p{n-1}^r$ choices: one of $n-1$ diagonal operators (that are
linearly independent of the identity) for each of $r$ tensor factors.
We discuss the exact choice of operators in more detail in Appendix
\ref{sec:operator_content}.  To summarize, we define $n$ single-spin
operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$, for
which $Z_\nu=\pm\sigma_\z$ in the case of SU(2), and we choose a list
of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
$\sum_\mu z_\mu=r$ to define the $r$-spin operator
\begin{align}
  X_z \equiv \bigotimes_{\nu=1}^n Z_\nu^{\otimes z_\nu}
  = \underbrace{Z_1 \otimes \cdots \otimes Z_1}_{z_1~\t{times}}
  \otimes
  \underbrace{Z_2 \otimes \cdots \otimes Z_2}_{z_2~\t{times}}
  \otimes \cdots \otimes
  \underbrace{Z_n \otimes \cdots \otimes Z_n}_{z_n~\t{times}}.
\end{align}
Our starting basis for $\M_r$ then consists of states
$\ket{m,\m w,X_z}\equiv\O\p{\m w,X_z}\ket{m}$, where we enforce that
$z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number
of the single-spin state $\nu$), and that $z$ is zero in an entry
where $m$ is maximal: $z_\nu=0$ for
$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  This second restriction is
equivalent to removing an operator from $\set{Z_\nu:\nu\in\ZZ_n}$,
making the remaining set linearly independent of the identity
operator.  The basis $\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally
complete, but not orthonormal, so as a final step we use the states
$\ket{m,\m w,X_z}$ to (numerically) construct an orthonormal basis
$\B_r$ for $\M_r$ (see Appendix \ref{sec:operator_content}).

@@

While the above restrictions on $z$ eliminate some trivial and
redundant states from the basis $\set{\ket{m,\m w,z}}$ for $\M_r$,
they do not guarantee that there will be {\it no} trivial states or
redundancies left over.  Furthermore, even after restrictions the
remaining basis is not orthogonal, which is to say that generally
$\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.  In order to
construct an orthonormal basis for $\M_r$, we therefore need to build
a matrix $B_m$ with matrix elements
\begin{align}
  B_m\p{\m v,y;\m w,z} \equiv \bk{m,\m v,y|m,\m w,z}
  = \bk{m| \O\p{\m v,X_y}^\dag \O\p{\m w,X_z} |m},
\end{align}
where $y,z$ satisfy appropriate restrictions, and $\m v,\m w$ are
chosen from the set of tensors that generate eigenstates of $H_0$.
Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form
$\sum_{\m w,z}c_{\m w z}\ket{m,\m w,z}$, whose squared norms are equal
to their corresponding eigenvalues.  Note that the matrix $B_m$ has a
block-diagonal structure, as it cannot (for example) couple states
$\ket{m,\m v,y},\ket{m,\m w,z}$ whose tensors $\m v,\m w$ generate
states with different interaction energies.  We therefore only need to
build and diagonalize this matrix independently within each block.
